Deuteron Cross Section Measurements up to $p_{\mathrm{r}}\sim1$ GeV/c Recoil Momentum

### 2020-05-07 MY_PUBLICATIONS/PRL/draft_v5_HCsuggestions/body.tex
updated

For each spectrometer, a logical hardware signal was created from  the coincidence of hits in three of the four scintillator planes. The event trigger was the the coincidence of these two signals. \\

@@

measured in this experiment. As a result, most of the virtual photon momentum is transferred to the proton which scatters at angles relative to $\vec{q}$ in the range $0.4^{\circ}\lesssim \theta_{pq}\lesssim21.4^{\circ}$.

@@

\indent The kinematics of the recoiling neutron was reconstructed using energy-momentum conservation. The recoil momentum is defined as $\vec{p_{\mathrm{r}}} = \vec{q} - \vec{p_{\mathrm{f}}}$
and the nuclear binding (or ``missing'') energy as $E_{\mathrm{m}} = \omega - T_{p} - T_{\mathrm{r}}$ where $\vec{p_{\mathrm{f}}}$ is the final proton momentum, $\vec{q}$ is the

@@

\indent For $^{1}\mathrm{H}(e,e'p)$, the corrected data yield was compared to SIMC calculations using J. Arrington's proton form factor parametrization \cite{PhysRevC.69.022201} to check the spectrometer acceptance

@@

\indent The systematic uncertainties due to our limited knowledge of the beam energy and spectrometer angle/momentum settings were determined point-to-point in ($\theta_{nq}, p_{\mathrm{r}}$) bins for each missing momentum setting,
and added in quadrature for overlapping $p_{r}$ bins. For $\theta_{nq}= 35^{\circ}, 45^{\circ}$ and $75^{\circ}$ (presented in this Letter) the overall kinematic uncertainty varied up to 6.5$\%$.

@@

\indent The data were radiatively corrected for each bin in ($\theta_{nq}, p_{\mathrm{r}}$) by multiplying the measured cross sections by the ratio of the calculated particle yield excluding and including radiative effects. The SIMC simulation code was used for these calculations with the Deuteron Model by J.M. Laget including FSI \cite{LAGET2005}.

@@

over the kinematic bin. The systematic uncertainty associated with the radiative and bin-centering corrections was investigated using the J.M. Laget PWIA and FSI models but negligible effects on the cross sections were found.\\
\indent The experimental and theoretical reduced cross sections were extracted and are defined as follows:

@@

@@

Within the PWIA, $\sigma_{\mathrm{red}}$ corresponds to the proton momentum distribution inside the deuteron. \\

@@

$p_{\mathrm{r}}$ for three recoil angle settings at $Q^{2}=4.5\pm0.5$ (GeV/c)$^{2}$. For the two highest settings ($p_{r}=580, 750$ MeV/c), a weighted average of the cross sections were taken in the overlapping regions of $p_{r}$ . The results from the previous experiment \cite{PhysRevLett.107.262501} at a $Q^{2}=3.5\pm0.25$ (GeV/c)$^{2}$ are plotted as well (red square). The good agreement between the Hall A and C data at lower $p_{r}$

@@

for the CD-Bonn (magenta) and AV18 (green) potentials were performed by M. Sargsian \cite{PhysRevC.82.014612} and those for the Paris potential  (blue) were  by J.M. Laget \cite{LAGET2005}.\\

@@

behaviour (smaller falloff) that overshadows any possibility of extracting the momentum distributions.\\

@@

\begin{figure}[t!]

@@

At $\theta_{nq}=45^{\circ}$ the FSI dominance starts earlier for all models above 800 MeV/c and for the CD-Bonn based calculation above 600 MeV/c. \\
\indent Overall, it is interesting to note that none of the calculations can reproduce the measured $p_r$ dependence above 600 MeV/c in a region where FSI are still relatively small $(<30\%)$.  This behavior of the data is new

@@

At $p_{\mathrm{r}}>300$ MeV/c, the data was statistically limited as our focus was on the smaller recoil angles. The Hall A data, however, shows a reasonable agreement with the FSI from all models which gives us confidence in our understanding

@@

very high neutron recoil momenta ($p_{r}>500$ MeV/c) at kinematics where FSI were expected to be small and the cross section was dominated by PWIA and sensitive to the
short range part of the deuteron wavefunction. The experimental reduced cross sections were extracted and found to be in good agreement with the Hall A data.

@@

momentum region ($p_{r}>500$ MeV/c) to better understand the large deviations observed between the different models and data.

@@

under grant No. NRC-HQ-84-14-G-0040 and the Doctoral Evidence Acquisition (DEA) Fellowship.

>>>>

For each spectrometer, a logical hardware signal was created from  the coincidence of hits in three of the four scintillator planes. The event trigger was the coincidence of these two signals. \\

@@

measured in this experiment. As a result, most of the virtual photon momentum is transferred to the proton, which scatters at angles relative to $\vec{q}$ in the range $0.4^{\circ}\lesssim \theta_{pq}\lesssim21.4^{\circ}$.

@@

\indent The kinematics of the recoiling neutron was reconstructed using energy-momentum conservation. The recoil momentum is defined as $\vec{p_{\mathrm{r}}} = \vec{q} - \vec{p_{\mathrm{f}}}$,
and the nuclear binding (or ``missing'') energy as $E_{\mathrm{m}} = \omega - T_{p} - T_{\mathrm{r}}$, where $\vec{p_{\mathrm{f}}}$ is the final proton momentum, $\vec{q}$ is the

@@

\indent For $^{1}\mathrm{H}(e,e'p)$, the corrected data yield was compared to SIMC calculations using Arrington's proton form factor parametrization \cite{PhysRevC.69.022201} to check the spectrometer acceptance

@@

%\indent The systematic uncertainties due to our finite knowledge of the beam energy and spectrometer angle/momentum settings were determined point-to-point in ($\theta_{nq}, p_{\mathrm{r}}$) bins for each missing momentum setting,
%and added in quadrature for overlapping $p_{\mathrm{r}}$ bins.
\indent The systematic uncertainties on the cross section due to systematic error on the absolute beam energy and spectrometer angle/momentum settings were
determined point-to-point in ($\theta_{nq}$, $p_{\mathrm{r}}$) bins for each missing momentum setting, and added in quadrature for overlapping $p_{\mathrm{r}}$ bins. 
For $\theta_{nq}= 35^{\circ}, 45^{\circ}$ and $75^{\circ}$ (presented in this Letter) the overall kinematic uncertainty varied up to 6.5$\%$.

@@

\indent The data were radiatively corrected for each bin in ($\theta_{nq}, p_{\mathrm{r}}$) by multiplying the measured cross sections by the ratio of the calculated particle yield excluding and including radiative effects. The SIMC simulation code was used for these calculations with the Deuteron Model by Laget including FSI \cite{LAGET2005}.

@@

over the kinematic bin. The systematic uncertainty associated with the radiative and bin-centering corrections was investigated using the Laget PWIA and 

@@

\noindent FSI models but negligible effects on the cross sections were found. The experimental and theoretical reduced cross sections were extracted and are defined as follows:

@@

Within the PWIA, $\sigma_{\mathrm{red}}$ corresponds to the PWIA cross section from the scattering of a proton in the deuteron. \\

@@

$p_{\mathrm{r}}$ for three recoil angle settings at $Q^{2}=4.5\pm0.5$ (GeV/c)$^{2}$. For the two highest settings ($p_{\mathrm{r}}=580, 750$ MeV/c), a weighted average of the cross sections were taken in the overlapping regions of $p_{\mathrm{r}}$ . The results from the previous experiment \cite{PhysRevLett.107.262501} at a $Q^{2}=3.5\pm0.25$ (GeV/c)$^{2}$ are plotted as well (red square). The good agreement between the Hall A and C data at lower $p_{\mathrm{r}}$

@@

for the CD-Bonn (magenta) and AV18 (green) potentials were performed by Sargsian \cite{PhysRevC.82.014612} and those for the Paris potential (blue) were by Laget \cite{LAGET2005}.\\

@@

behavior (smaller fall-off) that overshadows any possibility of extracting the momentum distributions.\\

@@

\begin{figure}[!t]

@@

At $\theta_{nq}=45^{\circ}$, the FSI dominance starts earlier for all models above 800 MeV/c and for the CD-Bonn based calculation above 600 MeV/c. \\
\indent Overall, it is interesting to note that none of the calculations can reproduce the measured $p_{\mathrm{r}}$ dependence above 600 MeV/c in a region where FSI are still relatively small $(<30\%)$.  This behavior of the data is new

@@

At $p_{\mathrm{r}}>300$ MeV/c, the data were statistically limited as our focus was on the smaller recoil angles. The Hall A data, however, shows a reasonable agreement with the FSI from all models which gives us confidence in our understanding

@@

very high neutron recoil momenta ($p_{\mathrm{r}}>500$ MeV/c) at kinematics where FSI were expected to be small and the cross section was dominated by PWIA and sensitive to the
short range part of the deuteron wave function. The experimental reduced cross sections were extracted and found to be in good agreement with the Hall A data.

@@

momentum region ($p_{\mathrm{r}}>500$ MeV/c) to better understand the large deviations observed between the different models and data.

@@

under grant No. NRC-HQ-84-14-G-0040, the Doctoral Evidence Acquisition (DEA) Fellowship and the Natural Sciences and Engineering Research Council of Canada (NSERC).

### 2020-05-07 MY_PUBLICATIONS/PRL/draft_v5_HCsuggestions/introduction.tex
updated

having to include explicit quark degree of freedoms, an issue of fundamental importance in nuclear physics \cite{sargsian_2015}. As of the present time, there are only a few nuclear physics experiments for

@@

from FSI, MEC and IC on the $^{2}\mathrm{H}(e,e'p)n$ cross-section and to determine the kinematics at which they are either suppressed (MEC and IC) or under control (FSI).  \\
\indent At large $Q^{2}$, FSI are described by the General Eikonal Approximation (GEA) \cite{Sargsian_2001,PhysRevC.56.1124,sargsian_2015} which predicts a strong dependence of FSI on neutron recoil angles $\theta_{nq}$.

@@

experiments determined that the cross-sections for fixed missing momenta indeed exhibited a strong angular dependence with neutron recoil angles, peaking

@@

\indent In Hall A, the pair of high resolution spectrometers (HRS) made it possible to measure the missing momentum dependence of the cross section for fixed neutron recoil angles ($\theta_{nq}$) reaching missing momenta up to $p_{\mathrm{r}}=550$ MeV/c at $Q^{2}=3.5\pm0.25$ (GeV/c)$^{2}$. For the first time very different momentum distributions were found for $\theta_{nq}=35\pm5^{\circ}$

@@

to $Q^{2}=4.5\pm0.5$ (GeV/c)$^{2}$ and neutron recoil momenta up to $p_{\mathrm{r}}\sim 1$ GeV/c which is almost double of the maximum recoil momentum measured in Hall A \cite{PhysRevLett.107.262501}.
Measurements at such large $Q^{2}$ and high missing momenta required scattered electrons to be detected at about 8.5 GeV/c which was only made possible with the newly commissioned Hall C Super High Momentum Spectrometer (SHMS).

>>>>

having to include explicit quark degrees of freedoms, an issue of fundamental importance in nuclear physics \cite{sargsian_2015}. As of the present time, there are only a few nuclear physics experiments for

@@

from FSI, MEC and IC on the $^{2}\mathrm{H}(e,e'p)n$ cross section and to determine the kinematics at which they are either suppressed (MEC and IC) or under control (FSI).  \\
\indent At large $Q^{2}$, FSI are described by the General Eikonal Approximation (GEA) \cite{Sargsian_2001,PhysRevC.56.1124,sargsian_2015}, which predicts a strong dependence of FSI on neutron recoil angles $\theta_{nq}$.

@@

experiments determined that the cross sections for fixed missing momenta indeed exhibited a strong angular dependence with neutron recoil angles, peaking

@@

\indent In Hall A, the pair of high resolution spectrometers (HRS) made it possible to measure the missing momentum dependence of the cross section for fixed neutron recoil angles ($\theta_{nq}$) reaching missing momenta up to $p_{\mathrm{r}}=550$ MeV/c at $Q^{2}=3.5\pm0.25$ (GeV/c)$^{2}$. For the first time, very different momentum distributions were found for $\theta_{nq}=35\pm5^{\circ}$

@@

to $Q^{2}=4.5\pm0.5$ (GeV/c)$^{2}$ and neutron recoil momenta up to $p_{\mathrm{r}}\sim 1$ GeV/c, which is almost double of the maximum recoil momentum measured in Hall A \cite{PhysRevLett.107.262501}.
Measurements at such large $Q^{2}$ and high missing momenta required scattered electrons to be detected at about 8.5 GeV/c, which was only made possible with the newly commissioned Hall C Super High Momentum Spectrometer (SHMS).

### 2020-05-07 MY_PUBLICATIONS/PRL/draft_v5_HCsuggestions/prl.tex
updated

% The following information is for internal review, please remove them for submission
\widetext
\leftline{Version 04 as of \today}
\leftline{Primary authors: Carlos Yero, Werner Boeglin, Mark Jones}
\leftline{To be submitted to PRL}
\leftline{Comment to {\tt cyero002@fiu.edu, \tt boeglinw@fiu.edu, \tt jones@jlab.org} by April 10, 2020}
\centerline{INTERNAL DOCUMENT -- NOT FOR PUBLIC DISTRIBUTION}

@@

%\title{First Measurements of the D(e,e'p)n Cross Section at \\ Very High Recoil Momenta and Large Q$^{2}$}
%\title{$^{2}$H$(e,e'p)n$ Cross Section Measurements for Neutron Recoil Momenta up to $p_{r}\sim1.01$ GeV/c at $Q^{2}=4.5\pm0.5$ (GeV/c)$^{2}$}
\title{Deuteron Electro-Disintegration at Very High Missing Momenta}
\input authorlist.tex

@@

\begin{abstract}
$^{2}\mathrm{H}(e,e'p)n$ cross sections have been measured at 4-momentum transfers of $-Q^{2} = 4.5 \pm 0.5$ (GeV/c)$^{2}$
reaching neutron recoil (missing) momenta up $p_{\mathrm{r}}\sim$ 1.0 GeV/c. The data have been obtained at fixed neutron recoil angles
$\theta_{nq} = 35^\circ$, $45^\circ$ and $75^{\circ}$  with respect to the 3-momentum
transfer $\vec q$. The new data agrees well with previous data which reached $p_{r}\sim500$ MeV/c. At $\theta_{nq} = 35^\circ$ and $45^\circ$, final state interactions (FSI),
meson exchange currents (MEC) and isobar configurations (IC) are
suppressed and the plane wave impulse approximation (PWIA) provides the
dominant cross section contribution. The new data are compared to recent
theoretical calculations, where a significant disagreement for
missing momenta $p_{r}>700$ MeV/c has been observed. 
\end{abstract}

>>>>

% The following information is for internal review, please remove them for submission
\widetext
\leftline{Version 05 as of \today}
\leftline{Primary authors: Carlos Yero, Werner Boeglin, Mark Jones}
\leftline{To be submitted to PRL}
\leftline{Comment to {\tt cyero002@fiu.edu, \tt boeglinw@fiu.edu, \tt jones@jlab.org} by May 01, 2020}
\centerline{INTERNAL DOCUMENT -- NOT FOR PUBLIC DISTRIBUTION}

@@

%\title{First Measurements of the D(e,e'p)n Cross Section at \\ Very High Recoil Momenta and Large Q$^{2}$}
%\title{$^{2}$H$(e,e'p)n$ Cross Section Measurements for Neutron Recoil Momenta up to $p_{r}\sim1.01$ GeV/c at $Q^{2}=4.5\pm0.5$ (GeV/c)$^{2}$}
%\title{Deuteron Electro-Disintegration up to $\sim$ 1 GeV/c Missing Momentum}
\title{Deuteron Cross Section Measurements up to $p_{\mathrm{r}}\sim1$ GeV/c Recoil Momentum}

\input authorlist.tex

@@

\begin{abstract}
$^{2}\mathrm{H}(e,e'p)n$ cross sections have been measured at 4-momentum transfers of $-Q^{2} = 4.5 \pm 0.5$ (GeV/c)$^{2}$
reaching neutron recoil (missing) momenta up $p_{\mathrm{r}}\sim$ 1.0 GeV/c. The data have been obtained at fixed neutron recoil angles
$\theta_{nq} = 35^\circ$, $45^\circ$ and $75^{\circ}$  with respect to the 3-momentum
transfer $\vec q$. The new data agrees well with previous data which reached $p_{\mathrm{r}}\sim500$ MeV/c. At $\theta_{nq} = 35^\circ$ and $45^\circ$, final state interactions (FSI),
meson exchange currents (MEC) and isobar configurations (IC) are
suppressed and the plane wave impulse approximation (PWIA) provides the
dominant cross section contribution. The new data are compared to recent
theoretical calculations, where a significant disagreement for
missing momenta $p_{\mathrm{r}}>700$ MeV/c has been observed. 
\end{abstract}
